---
title: 'Normal data: How much data is within 1 SD of the mean?'
author: "SMWadgymar, KGSmith, DMCharifson"
date: '2022-08-11'
output: html_document
---

This script will guide you through doing some simple calculations in R to test "the empirical rule". 

This rule of thumb states that,for normal distributions:

about 68% of the data will be within ±1 standard deviation (sd) of the mean value

about 95% of the data will be within ±2 sd of the mean.

Here, we will be exploring how the shape of a dataset affects this rule and our interpretation of sd.



####Normal data

First, we will create a dataset drawn randomly from a normal distribution.  Doing this will ensure that the data we are working with are pretty much perfectly normal, which means this is a theoretical test of the empirical rule.

Here is how we will simulate the data:

```{r}
norm.data=rnorm(10000)
```

The above creates a dataset of 10,000 values drawn from a normal distribution with a mean of 0 and an sd of 1 (the default settings in R).

Let's first explore the shape of the dataset to confirm that it's normal.  How would you describe the shape of the data? Use words other than 'normal'.


```{r}
hist(???, col="darkolivegreen4", breaks=50, xlab="# of sd from the mean")
```

Change the value of breaks lower and higher to see how it affects the graph.



Below, find the mean and SD of this data using mean() and sd().

```{r}

```

Q: Based on the distribution of the data, do you think the empirical rule will apply and that ~68% of data values will be within +/- 1 sd of the mean?  Why or why not?

A: 



We can actually find out how many datapoints are within +/- 1 sd of the mean using a few simple steps.

1) You know the mean and sd of the data.  From these values, define the lower and upper boundaries of an area that is 1 sd wide on either side of the mean value.  

```{r}
lower = mean(???) - sd(???)

upper = ???
```

2) Using these boundaries, create a subset that includes only the data in the middle of the range (i.e., +/- 1 sd from the mean).  There are several ways to do this, but here is one way that uses the subset() function. We have previously used brackets for subsetting (data[data=="type of observation"] or data[data<="numeric threshold"]).

```{r}
#Fill in the ??? with the appropriate code.
#the first argument is the data frame it is going to subset (norm.data). Then there is an argument for the criteria that it will be subset.
#Here there is a combination of criteria, it must fit the first criteria and the second, this is because of the use of the & symbol below.So it is specifying that the subset will consist of observations within norm.data that are both less than or equal to some value AND greater than or equal to some other value. These values are what you need to enter.

middle = subset(norm.data, norm.data <= ??? & norm.data >= ???)
```


3) Lastly, find out how many data values are within this middle range and then calculate what proportion of all data is within +/- 1 sd of the mean.

A useful function for this is length(x), where x is some variable. 

To solve this you must determine how many values are in the middle range, then calculate the proportion of values that fall within that middle range.

```{r}

```

Q: What proportion of data fell within +/- 1 sd of the mean?  What does this mean with respect to how well the empirical rule worked for this datset?  Why do you think that is?

A: 




####Human height data

Now you will do the same sort of analysis for another data set. This time we will not be simulating data, you will be working the cdc data set again. This means that now we will be dealing with a data frame instead of just a vector. So you might need to specify particular variables within that dataframe.

First, read the data in.

```{r}
source("http://www.openintro.org/stat/data/cdc.R")
```

Use the code from the first dataset above to answer the following questions about the human height dataset.

Look at the distribution of this data using a histogram.

```{r}

```

Find the mean and standard deviation of height using mean() and sd()

```{r}

```

Q: Do you think the "empirical rule" will apply and that 68% of data values will be within +/- 1 sd of the mean?  Why or why not?

A: 


Find out how many datapoints are within +/- 1 sd of the mean in three steps.

1) Define the boundaries of a zone that is 1 sd wide on either side of the mean value of height.

```{r}

```

2) Using these boundaries, create a subset that includes only the data in the middle of this range (i.e., +/- 1 sd from the mean).

```{r}
#Important note. Last time we were subsetting from a vector (list of values). Now we are subsetting from a data frame. this means that there are multiple columns each with different variables. You want to base your subset for the variable "height". You may still use a similar structure to the previous example in part 1, but you will need to do something to specify "height"


```

3) Lastly, find out how many data values are within this middle range and then calculate what proportion of all the data is within +/- 1 sd of the mean.

Note: There are two ways in which you may have subsetted the data in the previous step. One will result in a vector (list of values) the other will result in a new data frame with fewer rows. The function length() can tell you the number of observations in a vector, but if you ask length() to tell you about a data frame it will tell you the number of columns. You can do length(dataframe$variable) to correct this. Or you could do nrow(dataframe), which will tell you the number of rows in the data frame. Note this only applies when working with a dataframe, if you subset it to the extent that you are only working with a vector then length(vector) is fine!

```{r}

```

Q: What proportion of the data fell within +/- 1 sd of the mean?  What does this mean with respect to how well the empirical rule worked for this dataset?  Why do you think this is?

A:




##Transcript length data

This dataset contains transcript length data.

First, read the data in.

```{r}
coding.mrna=read.csv("https://docs.google.com/spreadsheets/d/1gnZefB6AQ2hSoZaSp4r40cxQ7vi4KIKnGDNkdCgAjKc/gviz/tq?tqx=out:csv&sheet=coding.mrna")
```


Let's create a datset that includes only the transcript length from Chromosome 2, which will make things a bit easier.  There are many different ways to subset data.  

```{r}
chr2.length=subset(coding.mrna, ???)
```


Use the code from the dataset above to answer the following questions about the transcript length data.

Look at the distribution of transcript length using a histogram.

```{r}

```

Find the mean and standard deviation of transcript length using mean() and sd()

```{r}

```

Q: Do you think the "empirical rule" will apply and that 68% of data values will be within +/- 1 sd of the mean?  Why or why not?

A: 


Find out how many datapoints are within +/- 1 sd of the mean in three steps.

1) Define the boundaries of a zone that is 1 sd wide on either side of the mean value of height.

```{r}

```

2) Using these boundaries, create a subset that includes only the data in the middle of this range (i.e., +/- 1 sd from the mean).

```{r}

```

3) Lastly, find out how many data values are within this middle range and then calculate what proportion of all the data is within +/- 1 sd of the mean.

Note: There are two ways in which you may have subsetted the data in the previous step. One will result in a vector (list of values) the other will result in a new data frame with fewer rows. The function length() can tell you the number of observations in a vector, but if you ask length() to tell you about a data frame it will tell you the number of columns. You can do length(dataframe$variable) to correct this. Or you could do nrow(dataframe), which will tell you the number of rows in the data frame. Note this only applies when working with a dataframe, if you subset it to the extent that you are only working with a vector then length(vector) is fine!

```{r}

```

Q: What proportion of the data fell within +/- 1 sd of the mean?  What does this mean with respect to how well the empirical rule worked for this dataset?  Why do you think this is?

A:


